\paperheader{2016 National Round}
\begin{problems}
    \problem The sequence $\log_{12}162$, $\log_{12}x$, $\log_{12}y$,
    $\log_{12}z$, $\log_{12}1250$ is an arithmetic progression. What is $x$?

    \problem A triangular array of 2016 coins has 1 coin in the first row, 2
    coins in the second row and 3 coins in the third row, and so on up to $N$
    coins in the $N$th row. What is the value of $N$?

    \problem Tu Tu cuts a circular paper disk of radius 12 cm along two radii
    to form two sectors, the smaller having a central angle of $120^\circ$. He
    makes two circular cones, using each sector to form the lateral surface of
    a cone. What is the ratio of the height of the smaller cone to that of the
    larger?

    \problem The sequence $u_{1}$, $u_{2}$, $u_{3}$, $\ldots$ has the property
    that every term beginning with the third is the sum of the previous two
    terms. That is, \[u_{n} = u_{n - 2} + u_{n - 1} \text{  for  } n \geq 3.\]
    Suppose that $u_{9} = 110$ and $u_{7} = 42$. What is $u_{4}$?

    \problem Line $\ell_{1}$ has equation $3x - 2y = 1$ and goes through $A =
    (-1, -2)$. Line $\ell_{2}$ has equation $y = 1$ and meets line $\ell_{1}$
    at point $B$. Line $\ell_{3}$ has positive slope, goes through point $A$
    and meets $l_{2}$ at point $C$. The area of $\triangle ABC$ is 3. What is
    the slope of line $\ell_{3}$?

    \problem In $\triangle ABC$, $AB = 13$, $BC = 14$ and $CA = 15$. Distinct
    points $D$ and $E$ lie on segments $BC$ and $CA$ respectively such that $AD
    \perp BC$ and $DE \perp AC$. Find the exact length of segment $DE$. 

    \problem In $\triangle ABC$, $AB = 6$, $BC = 7$ and $CA = 8$. Point $D$
    lies on $BC$ and $AD$ bisects $\angle BAC$. Point $E$ lies on $AC$ and $BE$
    bisects $\angle ABC$. The bisectors intersect at $F$. Find the ratios $AF :
    FD$ and $BF : FE$. 

    \problem The sum of an infinite geometric series is a positive number $S$,
    and the second term in the series is 1. What is the smallest possible value
    of $S$?

    \problem Prove that $\sqrt{ab} + \sqrt[3]{abc} \leq \frac{1}{3}(a + 4b +
    4c)$ for all positive numbers $a$, $b$ and $c$. 

    \problem Eight people are sitting around a circular table, each holding a
    fair coin. All eight people flip their coins and those who flip heads stand
    while those who flip tails remain seated. What is the number of ways that
    no two adjacent people will stand?

    \problem Let $ABCD$ be a cyclic quadrilateral and let the lines $CD$ and
    $BA$ meet at $E$. The line through $D$ which is tangent to the circle $ADE$
    meets the line $CB$ at $F$. Prove that $\triangle CDF$ is isosceles. 

    \problem Prove that $n^3 - n$ is divisible by 24, for all odd integers $n$. 

    \problem In how many ways can 7 paintings be posted on a wall,
    \begin{enumerate} \item if 3 of the paintings are always to appear
        altogether,
        
        \item if 3 of the paintings are never to appear altogether?
    \end{enumerate}

    \problem A box contains 2 red marbles, 2 green marbles, and 2 yellow
    marbles, Mg Gyi takes 2 marbles from the box at random; then Mg Latt takes
    2 of the remaining marbles at random; and then Mg Nge takes the last 2
    marbles. What is the probability that Mg Nge gets two marbles of the same
    colour. 

    \problem Prove by mathematical induction that \[1^3 + 2^3 + 3^3 + \cdots +
    n^3 =(1 + 2 + 3 + \cdots + n)^2\] for all positive integers $n$. 

    \problem If one of the roots of the equation: \[ax^3 + bx^2 + cx + d = 0\]
    is the geometric mean of the other two roots, prove that $ac^3 = b^3d$.
    
    \problem Let $f$ be a function such that \[x^2f(x) +
    f\left(\frac{x-1}{x}\right) = 2x^2,\] for all real numbers $x \ne 0$ and $x
    \ne 1$. Find the value of $f\left(\frac{1}{2}\right)$. 
\end{problems}
