\paperheader{2016 Regional Round}
\begin{problems}
    \problem In the sequence $u_{1}$, $u_{2}$, $\ldots$, $u_{n}$, $\ldots$, the
    $n$th term is defined by $u_{n} = 1 - \frac{1}{u_{n - 1}}$ for $n \geq 2$.
    If $u_{1} = 3$, compute $u_{2}$, $u_{3}$ and $u_{4}$. Write down
    $u_{2016}$.  
    
    \problem If $x$ and $y$ are positive integers such that $56 \leq x + y \leq
    59$ and $0.9 \leq \frac{x}{y} \leq 0.91$, find the value of $y^2 - x^2$. 
    
    \problem When 15 is added to a number $x$, it becomes the square of an
    integer. When 74 is subtracted from $x$, the result is a square of another
    integer. Find the number $x$. 
    
    \problem In a rectangle $ABCD$, the points $M$, $N$, $P$, $Q$ lie on $AB$,
    $BC$, $CD$ and $DA$ respectively, such that the areas of $\triangle AQM$,
    $\triangle BMN$, $\triangle CNP$ and $\triangle DPQ$ are equal. Prove that
    $MNPQ$ is a parallelogram.
    
    \problem Draw 6 circles in the plane such that every circle passes through
    exactly 3 centres of other circles. 
    
    \problem There are 2016 students in a secondary school. Every student
    writes a new year card. The cards are mixed up and randomly distributed to
    students. Suppose each student gets one and only one card. Find the
    expected number of students who get back their own cards. 
    
    \problem The points $P$, $A$, $B$ lie in that order on a circle with center
    $O$ such that $\angle POB < 180^\circ$. The point $Q$ lies inside the
    circle such that $\angle PAQ = 90^\circ$ and $PQ = BQ$. If $\angle AQB >
    \angle AQP$, prove that $\angle AQB - \angle AQP = \angle AOB$. 
    
    \problem A fair die is thrown three times. The results of the first, second
    and third throw are recorded as $x$, $y$ and $z$ respectively. Suppose that
    $x + y = z$. What is the probability that at least one of $x$, $y$ and $z$
    is 2?
    
    \problem An arithmetic progression and a harmonic progression have $a$ and
    $b$ for the first two terms. If their $n$th terms are $x$ and $y$
    respectively, show that $(x - a) : (y - a) = b : y$. 
    
    \problem Mr. Game owns 200 custom-made dice. Each die has four sides
    showing the number 2 and two sides showing the number 5. Mr. Game is about
    to throw all 200 dice together and find out the sum of all 200 results. How
    many possible values of this sum are there?
    
    \problem U Tet Toe wants to repair all 4 walls of his room. He has red
    paint, yellow paint and blue paint (which he cannot mix), and wants to
    paint his room so that adjacent walls are never of the same colour. In how
    many ways can U Tet Toe paint his room?
    
    \problem Given an equilateral triangle, what is the ratio of the area of
    its circumscribed circle to the area of the inscribed circle?
    
    \problem Solve the equation $\sqrt{3x^2 - 8x + 1} + \sqrt{9x^2 - 24x - 8} =
    3$. 
    
    \problem Prove by mathematical induction that \[1^2 + 3^2 + 5^2 + 7^2 +
    \cdots + (2n - 1)^2 = \frac{1}{3}n(4n^2 - 1).\] 
    
    \problem In how many ways can 7 identical T-shirts be divided among 4
    students, subject to the condition that each is to get at least 1 T-shirt. 
    
    \problem $\triangle ABC$ has $AB > AC$. A line $DEF$, equally inclined to
    $AB$ and $AC$, is drawn, meeting $AB$ at $F$, $AC$ at $E$ and $BC$ produced
    at $D$. Prove that $BD : DC = BF : CE$.
    
    (\textit{Hint:} Draw $BG \parallel CE$ meeting $DF$ produced at $G$.) 
    
    \problem The sum of the two smallest positive divisors of a positive
    integer $N$ is 6, while the sum of the largest positive divisors of $N$ is
    1122. Find $N$. 
    
    \problem If $p$, $q$ and $r$ are prime numbers such that $p < q < r$ and
    their product is 19 times their sum, find $p(q + r)$.
\end{problems}
