\paperheader{2017 Regional Round}
\begin{problems}
    \problem The four digit number $\overline{ABCD}$ is such that
    $\overline{ABCD} = A \times \overline{BCD} + \overline{ABC} \times D$. Find
    the smallest possible value of $\overline{ABCD}$. (Here $\overline{ABCD}$
    means $1000 \cdot A + 100 \cdot B + 10 \cdot C + D$.)
    
    \problem Find the remainder when 20172017201720172017201720172017 is
    divided by 72. 
    
    \problem Show that for all real numbers $x$, $y$, $z$ and $w$, \[\sin(x -
    w)\sin(y - z) + \sin(y - w)\sin(z - x) + \sin(z - w)\sin(x - y) = 0.\]
    
    \problem Two circles intersect at $A$ and $B$. A common tangent to the
    circles touches the circles at $P$ and $Q$. A circle is drawn through $P$,
    $Q$ and $A$, and the line $BA$ produced meets this circle again at $C$.
    Join $CP$ and $CQ$, and extend both to meet the given circles at $E$ and
    $F$ respectively. Prove that $P$, $Q$, $E$, $F$ are concyclic. 
    
    \problem If $x \geq 1$, prove that $x^3 - 5x^2 + 8x - 4 \geq 0$. 
    
    \problem Let $u_{1}, u_{2}, u_{3}, \ldots$ be a sequence of real numbers
    such that $u_{1} > 2$ and 
    \[u_{n + 1} = 1 + \frac{2}{u_{n}}\] 
    for $n \geq 1$. Prove that $u_{2n - 1} + u_{2n} > 4$ for all $n \geq 1$. 
    
    \problem Let $a$, $b$, $c$ be positive numbers such that $ab + bc + ca +
    abc = 4$. Prove that 
    \[\frac{1}{a + 2} + \frac{1}{b + 2} + \frac{1}{c + 2} = 1\]
    and
    \[a + b + c \geq 3.\]
    (You can use the following result: If $x$, $y$, $z$ are positive numbers,
    then $\frac{x + y + z}{3} \geq \sqrt[3]{xyz}$.)
    
    \problem On each side of a triangle, there are 4 distinct points other than
    triangle's vertices. Determine the number of triangles having the vertices
    at 3 of these 12 points. 
    
    \problem Two red dice and one blue die are thrown. What is the probability
    that the sum of the scores on the red dice is equal to the score on the
    blue die? 
    
    \problem $N$ is a two-digit number and $2N$ also has two digits. If $N$
    equals 2 times the sum of digits of $2N$, find all possible values of $N$.
    
    \problem In the figure, $AD \parallel BC$, $AB = BC = CD = \frac{1}{2}AD$
    and $\triangle APD$ is equilateral. $BP$ and $CP$ cut $AD$ at $Q$ and $R$
    respectively. If the area of $\triangle APD$ is 12, find the area of the
    trapezium $BQRC$.
    \begin{center}
        \begin{asy}
            import olympiad;
            size(5cm);
            defaultpen(fontsize(11pt));
            pen mydash = linetype(new real[] {5,5});
            pair A = dir(180);
            pair B = dir(240);
            pair C = dir(300);
            pair D = dir(0);
            pair T = extension(A, B, C, D);
            pair M = (0, 0);
            pair P = 2*M - T;
            pair Q = extension(P, B, A, D);
            pair R = extension(P, C, A, D);
            draw(A--B--C--D--cycle);
            draw(A--P--D);
            draw(B--P--C);
            dot("$A$", A, dir(A));
            dot("$B$", B, dir(B));
            dot("$C$", C, dir(C));
            dot("$D$", D, dir(D));
            dot("$P$", P, dir(90));
            dot("$Q$", Q, dir(135));
            dot("$R$", R, dir(45));
        \end{asy}
    \end{center}
    
    \problem
    \begin{enumerate}
        \item Show that $\frac{1}{1 \times 2} + \frac{1}{2 \times 3} +
            \frac{1}{3 \times 4} + \cdots + \frac{1}{(n - 1)n} = \frac{n -
            1}{n}$ for $n \geq 2$. 
        
        \item Show that for every integer $n \geq 2$, there exist positive
            integers $x_{1}, x_{2}, x_{3}, \ldots, x_{n}$ so that 
            \[ \frac{1}{x_{1}} + \frac{1}{x_{2}} + \frac{1}{x_{3}} + \cdots +
            \frac{1}{x_{n}} = 1.\] 
    \end{enumerate}
    
    \problem Let $a$, $b$, $c$, $d$, $e$ be five prime numbers forming an
    arithmetic progression with a common difference of 6. Find the smallest
    possible value of $a + b + c + d + e$. 
    
    \problem Function $f:\mathbb{R} \rightarrow \mathbb{R}$ satisfies 
    \[(a - b)f(a + b) + (b - c)f(b + c) + (c - a)f(c + a) = 0,\] 
    for all $a, b, c \in \mathbb{R}$. For $x \in \mathbb{R}$, letting $a =
    \frac{1}{2}(x - 1)$, $b = \frac{1}{2}(x + 1)$ and $c = \frac{1}{2}(1 - x)$,
    show that $f(x) = Ax + B$ where $A = f(1) - f(0)$ and $B = f(0)$.
    Conversely, show that the function $f:\mathbb{R} \rightarrow \mathbb{R}$
    defined by $f(x) = Ax + B$, where $A$ and $B$ are constants, satisfies the
    given equation. 
    
    \problem In a trapezium $ABCD$, with $AB \parallel CD$, there are two
    circles with diameter $AD$ and $BC$ respectively. Two circles do not
    intersect each other. Let $X$ and $Y$ which do not lie in $ABCD$ be two
    points on each of the circles. Show that 
    \[XY \leq \frac{1}{2}(AD + AB + DC + BC).\]
    
    \problem If $\alpha$, $\beta$ and $\gamma$ are roots of the equation $x^3 +
    ax^2 + bx + c = 0$, evaluate $\alpha^2 + \beta^2 + \gamma^2$ and $\alpha^3
    + \beta^3 + \gamma^3$. 
    
    \problem Prove by mathematical induction, that $3^{2n} + 7$ is divisible by
    8 for all $n \geq 1$. 
    
    \problem Each vertex of convex polygon $ABCDE$ is to be assigned a colour.
    There are 6 colours and the ends of each diagonal must have different
    colours. How many different colourings are possible?
\end{problems}
