\paperheader{2018 National Round}
\begin{problems}
    \problem To construct a triangle $ABC$, only given that $AB = 10$ and
    $\angle ABC$ is $30^\circ$. Find all values of $AC$ for which 
    \begin{enumerate}
        \item there are two possible triangles $ABC$, 
        
        \item there is only one triangle $ABC$, 
        
        \item there is no triangle $ABC$. 
    \end{enumerate}
    
    \problem In the given figure, $O$ is the center of the circle. $OA$ and
    $OC$ are radii with $OA = OC = 5$ units. Find the area of region $ABCD$,
    bounded by arc $ABC$, line segments $CD$ and $DA$, in terms of $\theta$, if
    the line segment $DA = 8$ units and $\angle AOC$ is $\theta^\circ$. 
    \begin{center}
        \begin{asy}
            import olympiad;
            size(5cm);
            defaultpen(fontsize(11pt));
            pen mydash = linetype(new real[] {5,5});
            usepackage("contour", "outline");
            texpreamble("\contourlength{1pt}");
            real s = 35;
            pair A = dir(s);
            pair C = dir(-s);
            pair B = dir(0);
            pair D = (3/5)*dir(180+s);
            pair O = (0, 0);
            pair H = foot(C, A, D);
            draw(A--O--C--cycle);
            draw(O--D--C);
            draw(circle(O, 1));
            draw(anglemark(C, O, A, 5));
            dot("$A$", A, dir(A));
            dot("$C$", C, dir(C));
            dot("$D$", D, dir(210));
            dot("$O$", O, dir(135));
            label("$B$", B, dir(0));
            label("\contour{white}{$\theta^\circ$}", 0.15*dir(0), dir(0));
        \end{asy}
    \end{center}
    
    \problem Prove that
    \begin{enumerate}
        \item $\cos 20^\circ$,
        
        \item $\log 21$, and 
        
        \item $\sqrt{3}$ 
    \end{enumerate}
    are irrational.
    
    \problem Triangle $ABC$ with $AB = c$, $BC = a$ and $CA = b$ is inscribed
    in a circle. Find the radius of the circle in terms of $a$, $b$ and $c$. 
    
    \problem Prove the following Ptolemy's theorem: \par In a cyclic
    quadrilateral $PQRS$, $$PQ \cdot SR + PS \cdot QR = PR \cdot SQ,$$ That is,
    the sum of the products of the opposite sides is equal to the product of
    the diagonals. 
    
    \problem $f(x)$ is a real-valued function defined on $a < x < b$ such that 
    \[f \left( \frac{x_{1} + x_{2}}{2} \right) \leq \frac{1}{2}(f(x_{1}) +
    f(x_{2})),\] 
    for all $a < x_{1}, x_{2} < b$. Prove that 
    \[f\left(\frac{x_{1} + x_{2} + \cdots + x_{n}}{n}\right)\leq
    \frac{1}{n}(f(x_{1}) + f(x_{2}) + \cdots + f(x_{n}))\] 
    for all $a < x_{1}, x_{2}, \ldots, x_{n} < b$. 
    
    \problem Point $A$ has position vector $\overrightarrow{OA}$ and point $B$
    has position vector $\overrightarrow{OB}$. Prove that for $0\leq \lambda
    \leq 1$, the vector $\lambda \cdot \overrightarrow{OA} + (1 - \lambda)
    \cdot \overrightarrow{OB}$ is a position vector of a point on the line
    segment $AB$. Prove also that any point on the line segment $AB$ has
    position vector $\lambda \cdot \overrightarrow{OA} + (1 - \lambda) \cdot
    \overrightarrow{OB}$ for some $0\leq \lambda \leq 1$. 
    
    \problem $f(x)$ is a real-valued function defined on the interval $a < x <
    b$ such that 
    \[f(tx_{1} + (1 - t)x_{2}) \leq tf(x_{1}) + (1 - t)f(x_{2})\] 
    for all $a < x_{1}, x_{2} < b$ and for all $0 \leq t \leq 1$. Prove that
    for each triple $x_1$, $x_2$, $x_3$ of distinct numbers in the interval, 
    \[\frac{(x_{3} - x_{2})f(x_{1}) + (x_{2} - x_{1})f(x_{3}) + (x_{1} -
    x_{3})f(x_{2})}{(x_{1} - x_{2})(x_{2} - x_{3})(x_{3} - x_{1})}\geq 0.\]
    
    \problem Equilateral triangle $ABC$ is inscribed in a circle. $P$ is a
    point on arc $AB$. Prove that $PA + PB = PC$. 
    
    \problem $\triangle ABC$ is inscribed in a circle. The tangent at $C$ and
    the line through $B$ parallel to $AC$ meet at $D$. The tangent at $B$ and
    the line through $C$ parallel to $AB$ meet at $E$. Prove that $BC^2 = BE
    \cdot CD$. 
    
    \problem $\triangle ABC$ is isosceles with $AB = AC$. $D$ is the midpoint
    of $BC$. $AB$ and $AC$ are produced to $X$ and $Y$ respectively such that
    $\angle XDY = \angle DCY$. Prove that $\triangle XBD$, $\triangle XDY$ and
    $\triangle DCY$ are similar triangles. 
    
    \problem $a_{1}$, $b_{1}$, $c_{1}$, $a_{2}$, $b_{2}$, $c_{2}$ are positive
    real numbers with $a_{1}c_{1} - b_{1}^2 \geq 0$ and $a_{2}c_{2} - b_{2}^2
    \geq 0$. Show that $(a_{1} + a_{2})(c_{1} + c_{2}) - (b_{1} + b_{2})^2 \geq
    0$. 
    
    \problem How many permutations of the word `TRIANGLE' have none of the
    vowels together?
    
    \problem From the group of $2n + 1$ people, how many ways to choose a group
    of $n$ people or less?
    
    \problem Show that the product of any positive integer and its $k - 1$
    successors is divisible by $k!$. 
    
    \problem How many integers between 1 and 1,000,000 have the sum of digits
    equal to 10?
    
    \problem Each side and diagonal of a regular hexagon is coloured either red
    or blue. Show that there is a triangle with all three sides of the same
    colour. 
    
    \problem Find the value of $2018^2 - 2017^2 + 2016^2 - 2015^2 + \cdots +
    2^2 - 1$.
\end{problems}
